Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load

### 2024-01-03 Paper/paper.tex
working through sonia edits

% journal

@@

\usepackage[acronym]{glossaries}
\usepackage{hhline}
\usepackage{threeparttable} % to use table notes
%\usepackage[draft]{changes} % usage: \deleted{old} \added{new} \replaced{new}{old} also \hl{} for highlighted text
%\usepackage[final]{changes}

@@

% MDPI internal command: Title for citation in the left column
\TitleCitation{Optimal Peak Shaving With Storage Battery and Bifacial Solar PV}

@@

Even in cases where the electric load isn't growing, large solar photovoltaic (PV) arrays relative to the load will create relative peaks and dips in the net load seen by the network as clouds pass overhead. As the solar ramp-down rate due to clouds can be much faster than the ramp-up rate of many traditional thermal generators, grid operators will need to predict and anticipate these relative peaks. With an all-sky camera and machine vision algorithms, 15-minute nowcasting cloud identification may only reach 84.2\% accuracy \cite{Nespoli2022}. The remaining peaks will likely need to be managed by battery or other energy storage systems with extremely fast power ramp rates.

@@

Peak load management, or peak shaving, is the process of reducing the maximum load at a given point in an electrical network. This will necessarily increase the load factor ($P_{load,avg} P_{load,max}^{-1}$) even if load curtailment is the only method used. Other methods include load scheduling, load shifting using energy storage, and reduction of the effective load with distributed generation (DG) \cite{Uddin2018}. Peak shaving may include the concept of a peak load threshold which defines a maximum desired peak load during a period, typically on the order of hours. Grid operators may use peak shaving to increase power quality, defer future infrastructure upgrades, decrease losses, or reduce energy costs. End-users typically reduce their peaks solely on a cost basis. Lastly, some carbon reduction is possible by reducing the on/off transitions of peaker plants due to load peaks \cite{Uddin2018}.

@@

\begin{table}
  \centering
  \caption{Economic and Technical Peak Shaving Comparison}
  \label{tab:econ-tech-peak-shaving}
  \begin{tabularx}{\linewidth}{X X X}
    \toprule
    Feature                                      & Technical Peak Shaving                    & Economic Peak Shaving             \\
    \midrule
    Consequence of exceeding the power threshold & Damage to infrastructure, possible outage & Monetary cost depending on tariff \\
    Time resolution                              & $\mu s$ to 10 s                           & 15-60 min (typical)               \\
    Valid times of day                           & All                                       & Limited (e.g. 16:00 to 21:00)     \\
    Peak is recalculated every..                 & Continuous (never recalculated)           & Month (also: day, year)           \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

  In the literature, peak shaving strategies that dispatch a battery system often don't consider the retail cost of energy and rarely use an optimized threshold approach or vertical bifacial solar PV. In \cite{Reihani2016} authors forecast load and choose a state of charge (SOC) trajectory to manage load peaks. Solar power smoothing is applied in works such as \cite{Lavrova2012}. Linear programming optimization in \cite{Lee2014} reduces the on-peak load of residential EV charging by dispatching a battery. In \cite{Gazafroudi2018} a multi-agent system buys and sells energy in a local energy market for a residence with solar PV and a dispatchable vehicle-to-home EV battery. In \cite{Zheng2015} authors optimize peak shaving based on a rate tariff and a technical power threshold. Distributed solar PV  helps reduce net load and tracking arrays produce more energy in the late afternoon \cite{Faranda2011} but the increase may be reduced by common backtracking methods \cite{Dolara2012}. Vertical bifacial has been benchmarked in the literature and evaluated for new applications such as public highway solar installations \cite{Niccolai2022}. Vertical bifacial solar is coupled with a storage battery in \cite{Wallberg2022} and in \cite{Castellucci2022} with a rule-based peak shaving algorithm, reducing EV charging load peaks by 38.6\% and 79\%, respectively. However, the retail energy cost isn't estimated and a comparison is not made to understand the relative merit of vertical bifacial compared to south-facing modules.

  The following paper is organized into the following sections: \ref{methodology}, \ref{case-studies}, \ref{results}, \ref{conclusion}, and \ref{bibliography}.

  %\hypertarget{methodology}{%
  \section{Methodology}\label{methodology}%}

@@

\begin{table}[!h]
  \centering
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $[kW_{ac}]$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $[W/m^2]$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $[kW_{ac}]$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[!h]
  \centering
  \caption{Retail Electric Tariff. An applicable electric tariff schedule with power and energy prices for several different TOU periods. From California PG\&E.}
  \label{tab:tariff}
  \begin{tabularx}{\textwidth}{XXXXX}
    \toprule
    Name            & TOU Period                         & Summer (Jun 1 - Sept 30) $[\$]$ & Winter (Oct 1 -Feb 28) $[\$]$ & Spring (Mar 1 - May 30) $[\$]$ \\
    \midrule
    All hours       & 0:00-0:00     3                    & 26.07/$kW$                      & 26.07/$kW$                    & 26.07/$kW$                     \\
    Super off-peak  & 9:00-14:00                         & -\/-                            & -\/-                          & 0.079/$kWh$                    \\
    Off-peak spring & 0:00-9:00, 14:00-16:00, 21:00-0:00 & -\/-                            & -\/-                          & 0.132/$kWh$                    \\
    Off-peak winter & 0:00-16:00, 21:00-0:00             & -\/-                            & 0.132/$kWh$                   & -\/-                           \\
    Off-peak summer & 0:00-14:00, 23:00-0:00             & 0.132/$kWh$                     & -\/-                          & -\/-                           \\
    Partial-peak    & 14:00-16:00, 21:00-23:00           & 6.81/$kW$, 0.159/$kWh$          & -\/-                          & -\/-                           \\
    Peak            & 16:00-21:00                        & 32.90/$kW$, 0.196/$kWh$         & 2.22/$kW$, 0.172/$kWh$        & 2.22/$kW$, 0.172/$kWh$         \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}
  \centering
  \caption{Case Studies. All modules are the Prism 350 W bifacial. The base case array faces South and is tilted to 20°. The following four cases orient an increasing percentage of the modules West and tilt to 90°, with an expected linear decrease in total energy and solar yield.}
  \label{tab:casestudies}
  \begin{tabularx}{\textwidth}{XXXXXX}
    \toprule
    Case Study                     & Modules & Orientation & Tilt & Total Energy ($MWh$)   & Solar Yield ($\frac{kWh}{kW_{DC}}$) \\
    \midrule
    South 20° (base)               & 368     & South       & 20°  & 151.7                  & 1180.8                              \\
    \hline
    \multirow{2}{*}{25\% West 90°} & 276     & South       & 20°  & \multirow{2}{*}{146.5} & \multirow{2}{*}{1140.4}             \\
                                   & 92      & West        & 90°  &                        &                                     \\
    \hline
    \multirow{2}{*}{50\% West 90°} & 184     & South       & 20°  & \multirow{2}{*}{141.1} & \multirow{2}{*}{1098.2}             \\
                                   & 184     & West        & 90°  &                        &                                     \\
    \hline
    \multirow{2}{*}{75\% West 90°} & 92      & South       & 20°  & \multirow{2}{*}{136.1} & \multirow{2}{*}{1059.6}             \\
                                   & 276     & West        & 90°  &                        &                                     \\
    \hline
    100\% West 90°                 & 368     & West        & 90°  & 131.0                  & 1019.3                              \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Total Retail Electric Cost. Simulated total retail electric cost for three of the case studies (some omitted for conciseness) over the 10 months of available data and a given battery capacity. Cost reduction is calculated with respect to the South 20° base case. The best reduction in total retail cost is \$1422 for the 50\% West 90° case.}
  \label{tab:total-cost}
  \begin{tabularx}{\textwidth}{XXXXXXXX}
    \toprule
    Battery $[kWh]$ & \multicolumn{3}{l}{Cost $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\%]$}                                                                   \\
                    & South 20°                       & 50\% West 90°                             & 100\% West 90°                            & 50\% West 90° & 100\% West 90° & 50\% West 90° & 100\% West 90° \\
    \midrule
    25              & 31314                           & 30391                                     & 31730                                     & 923           & -416           & 3.0           & -1.3           \\
    50              & 27132                           & 25773                                     & 27022                                     & 1359          & 110            & 5.0           & 0.4            \\
    75              & 23893                           & 22775                                     & 23625                                     & 1118          & 268            & 4.7           & 1.1            \\
    100             & 21229                           & 19807                                     & 20804                                     & \textbf{1422} & 425            & 6.7           & 2.0            \\
    125             & 18954                           & 17623                                     & 18564                                     & 1331          & 390            & 7.0           & 2.1            \\
    150             & 17094                           & 15886                                     & 17286                                     & 1208          & -192           & 7.1           & -1.1           \\
    200             & 14547                           & 13689                                     & 15282                                     & 858           & -735           & 5.9           & -5.0           \\
    400             & 10730                           & 10969                                     & 12131                                     & -239          & -1401          & -2.2          & -13.1          \\
    600             & 8859                            & 9140                                      & 10333                                     & -281          & -1474          & -3.2          & -16.6          \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Solar capacity sensitivity analysis based on reduction in total retail cost relative to the base South 20° case. While the 50\% West 90° array continues to achieve the highest cost reduction, the battery capacity increases from 25 kWh to 125 kWh.}
  \label{tab:solar-sensitivity}
  \begin{tabularx}{\textwidth}{XXXX}
    \toprule
    Solar Capacity {[}\(kW_{DC}\){]} & Best Case for Cost Reduction         & Best
    Cost Reduction {[}\${]}          & Best Cost Reduction Battery Capacity
    {[}kWh{]}                                                                            \\
    \midrule
    64.4 (0.5x)                      & 50\% West 90°                        & 260  & 25  \\
    128.8 (1.0x)                     & 50\% West 90°                        & 1422 & 100 \\
    193.2 (1.5x)                     & 50\% West 90°                        & 2211 & 125 \\
    257.6 (2.0x)                     & 75\% West 90°                        & 2579 & 125 \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

%\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}%}

@@

Confidence in the results presented is medium-high given the use of true EV charging measurements, a robust Newton-Raphson gradient descent optimizer with verified optimality, a true electric tariff appropriate for the load site and solar data location, and sensitivity analyses performed on two variables. Future work should include more load data sets with more months of data, a more comprehensive battery model, an operational peak shaving strategy based on a load forecast, and testing at the Politecnico di Milano M2GLab microgrid.

%\hypertarget{bibliography}{%
\section{Bibliography}\label{bibliography}%}

>>>>

%journal

@@

%\usepackage[acronym]{glossaries}
%\usepackage{hhline}
%\usepackage{threeparttable} % to use table notes
\usepackage[draft]{changes} % usage: \deleted{old} \added{new} \replaced{new}{old} also \hl{} for highlighted text
%\usepackage[final]{changes}

@@

% MDPI internal command: Title for citation in the left column
\TitleCitation{Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load}

@@

Even in cases where the electric load isn't growing, solar photovoltaic (PV) arrays that are large relative to the load will create relative peaks and valleys in the load seen by the network as clouds pass overhead. As the solar ramp-down rate due to clouds can be much faster than the ramp-up rate of many traditional thermal generators, grid operators will need to predict and anticipate these relative peaks. With an all-sky camera and machine vision algorithms, 15-minute nowcasting cloud identification may only reach 84.2\% accuracy \cite{Nespoli2022}. The remaining peaks will likely need to be managed by other means.

@@

Peak load management, or peak shaving, is the process of reducing the maximum load at a given point in an electrical network. This will necessarily increase the load factor ($P_{load,avg} P_{load,max}^{-1}$) even if load curtailment is the only method used. Other methods include load scheduling, load shifting using energy storage (battery or otherwise), and reduction of the effective load with distributed generation (DG) \cite{Uddin2018}. Peak shaving may include the concept of a peak load threshold which defines a maximum desired peak load during a period, typically on the order of hours. Grid operators may use peak shaving to increase power quality, defer future infrastructure upgrades, decrease losses, or reduce energy costs. End-users typically reduce their peaks solely on a cost basis. Lastly, some carbon reduction is possible by reducing the on/off transitions of peaker plants due to load peaks \cite{Uddin2018}.

@@

\begin{table}
  \caption{Economic and Technical Peak Shaving Comparison}
  \label{tab:econ-tech-peak-shaving}
  \begin{tabularx}{\linewidth}{X X X}
    \toprule
    Feature                                      & Technical Peak Shaving                    & Economic Peak Shaving             \\
    \midrule
    Consequence of exceeding the power threshold & Damage to infrastructure, possible outage & Monetary cost depending on tariff \\
    Time resolution                              & $\mu s$ to 10 s                           & 15-60 min (typical)               \\
    Valid times of day                           & All                                       & Limited (e.g. 16:00 to 21:00)     \\
    Peak is recalculated every..                 & Continuous (never recalculated)           & Month (also: day, year)           \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

In the literature, peak shaving papers often don't consider the retail cost of energy and rarely utilize vertical bifacial solar PV or a threshold-based approach. In \cite{Reihani2016} authors forecast load and choose a state of charge (SOC) trajectory to manage load peaks. Linear programming optimization in \cite{Lee2014} reduces the on-peak load of residential EV charging by dispatching a battery. In \cite{Gazafroudi2018} a multi-agent system buys and sells energy in a local energy market for a residence with solar PV and a dispatchable vehicle-to-home EV battery. In \cite{Zheng2015} authors optimize peak shaving based on a rate tariff and a technical power threshold. Vertical bifacial has been benchmarked in the literature and evaluated for new applications such as public highway solar installations \cite{Niccolai2022}. Vertical bifacial solar is coupled with a storage battery in \cite{Wallberg2022} and in \cite{Castellucci2022} with a rule-based peak shaving algorithm, reducing EV charging load peaks by 38.6\% and 79\%, respectively. However, the retail energy cost isn't estimated and a comparison is not made to understand the relative merit of vertical bifacial compared to south-facing modules. Bifacial PV has been studied extensively in the literature, performing around 11\% better than comparable monofacial PV \cite{Ogliari2023}, .

\hl{Considering the above, this work builds on the existing body of knowledge by uniquely considering the benefit of bifacial solar in a peak shaving application and does so with an optimal threshold power approach. The objectives are a reproducible methodology to quantify that benefit and to explore the system dynamics to inform future investigations. The rest of the paper is organized as follows.}

\begin{enumerate}
	\item Methodology: Formal definition of the cost function and how the minimization problem is solved
	\item Case Studies: Five specific contexts of the problem are solved, each with a different solar array configuration
	\item Results: The minimal costs are evaluated for each case and compared to the case base, including sensitivity analysis
	\item Conclusions: The important findings are summarized and future directions explored
\end{enumerate}

\section{Methodology}\label{methodology}%}

@@

\begin{table}[!h]
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $[kW_{ac}]$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $[W/m^2]$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $[kW_{ac}]$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[!h]
  \caption{Retail Electric Tariff. An applicable electric tariff schedule with power and energy prices for several different TOU periods. From California PG\&E.}
  \label{tab:tariff}
  \begin{tabularx}{\textwidth}{XXXXX}
    \toprule
    Name            & TOU Period                         & Summer (Jun 1 - Sept 30) $[\$]$ & Winter (Oct 1 -Feb 28) $[\$]$ & Spring (Mar 1 - May 30) $[\$]$ \\
    \midrule
    All hours       & 0:00-0:00     3                    & 26.07/$kW$                      & 26.07/$kW$                    & 26.07/$kW$                     \\
    Super off-peak  & 9:00-14:00                         & -\/-                            & -\/-                          & 0.079/$kWh$                    \\
    Off-peak spring & 0:00-9:00, 14:00-16:00, 21:00-0:00 & -\/-                            & -\/-                          & 0.132/$kWh$                    \\
    Off-peak winter & 0:00-16:00, 21:00-0:00             & -\/-                            & 0.132/$kWh$                   & -\/-                           \\
    Off-peak summer & 0:00-14:00, 23:00-0:00             & 0.132/$kWh$                     & -\/-                          & -\/-                           \\
    Partial-peak    & 14:00-16:00, 21:00-23:00           & 6.81/$kW$, 0.159/$kWh$          & -\/-                          & -\/-                           \\
    Peak            & 16:00-21:00                        & 32.90/$kW$, 0.196/$kWh$         & 2.22/$kW$, 0.172/$kWh$        & 2.22/$kW$, 0.172/$kWh$         \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}
  \caption{Case Studies. All modules are the Prism 350 W bifacial. The base case array faces South and is tilted to 20°. The following four cases orient an increasing percentage of the modules West and tilt to 90°, with an expected linear decrease in total energy and solar yield.}
  \label{tab:casestudies}
  \begin{tabularx}{\textwidth}{XXXXXX}
    \toprule
    Case Study                     & Modules & Orientation & Tilt & Total Energy ($MWh$)   & Solar Yield ($\frac{kWh}{kW_{DC}}$) \\
    \midrule
    South 20° (base)               & 368     & South       & 20°  & 151.7                  & 1180.8                              \\
    \hline
    \multirow{2}{*}{25\% West 90°} & 276     & South       & 20°  & \multirow{2}{*}{146.5} & \multirow{2}{*}{1140.4}             \\
                                   & 92      & West        & 90°  &                        &                                     \\
    \hline
    \multirow{2}{*}{50\% West 90°} & 184     & South       & 20°  & \multirow{2}{*}{141.1} & \multirow{2}{*}{1098.2}             \\
                                   & 184     & West        & 90°  &                        &                                     \\
    \hline
    \multirow{2}{*}{75\% West 90°} & 92      & South       & 20°  & \multirow{2}{*}{136.1} & \multirow{2}{*}{1059.6}             \\
                                   & 276     & West        & 90°  &                        &                                     \\
    \hline
    100\% West 90°                 & 368     & West        & 90°  & 131.0                  & 1019.3                              \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Total Retail Electric Cost. Simulated total retail electric cost for three of the case studies (some omitted for conciseness) over the 10 months of available data and a given battery capacity. Cost reduction is calculated with respect to the South 20° base case. The best reduction in total retail cost is \$1422 for the 50\% West 90° case.}
  \label{tab:total-cost}
  \begin{tabularx}{\textwidth}{XXXXXXXX}
    \toprule
    Battery $[kWh]$ & \multicolumn{3}{l}{Cost $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\%]$}                                                                   \\
                    & South 20°                       & 50\% West 90°                             & 100\% West 90°                            & 50\% West 90° & 100\% West 90° & 50\% West 90° & 100\% West 90° \\
    \midrule
    25              & 31314                           & 30391                                     & 31730                                     & 923           & -416           & 3.0           & -1.3           \\
    50              & 27132                           & 25773                                     & 27022                                     & 1359          & 110            & 5.0           & 0.4            \\
    75              & 23893                           & 22775                                     & 23625                                     & 1118          & 268            & 4.7           & 1.1            \\
    100             & 21229                           & 19807                                     & 20804                                     & \textbf{1422} & 425            & 6.7           & 2.0            \\
    125             & 18954                           & 17623                                     & 18564                                     & 1331          & 390            & 7.0           & 2.1            \\
    150             & 17094                           & 15886                                     & 17286                                     & 1208          & -192           & 7.1           & -1.1           \\
    200             & 14547                           & 13689                                     & 15282                                     & 858           & -735           & 5.9           & -5.0           \\
    400             & 10730                           & 10969                                     & 12131                                     & -239          & -1401          & -2.2          & -13.1          \\
    600             & 8859                            & 9140                                      & 10333                                     & -281          & -1474          & -3.2          & -16.6          \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Solar capacity sensitivity analysis based on reduction in total retail cost relative to the base South 20° case. While the 50\% West 90° array continues to achieve the highest cost reduction, the battery capacity increases from 25 kWh to 125 kWh.}
  \label{tab:solar-sensitivity}
  \begin{tabularx}{\textwidth}{XXXX}
    \toprule
    Solar Capacity {[}\(kW_{DC}\){]} & Best Case for Cost Reduction         & Best
    Cost Reduction {[}\${]}          & Best Cost Reduction Battery Capacity
    {[}kWh{]}                                                                            \\
    \midrule
    64.4 (0.5x)                      & 50\% West 90°                        & 260  & 25  \\
    128.8 (1.0x)                     & 50\% West 90°                        & 1422 & 100 \\
    193.2 (1.5x)                     & 50\% West 90°                        & 2211 & 125 \\
    257.6 (2.0x)                     & 75\% West 90°                        & 2579 & 125 \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\section{Conclusions}\label{conclusions}

@@

Confidence in the results presented is medium-high given the use of true EV charging measurements, a robust Newton-Raphson gradient descent optimizer with verified optimality, a true electric tariff appropriate for the load site and solar data location, and sensitivity analyses performed on two variables.

Future work should expand the load case studies including multiple years and true solar production measurements, \hl{improve the battery state of charge estimation such as in} \cite{Eleftheriadis2023}, \hl{perform simulations using forecasting load such as in} \cite{Wood2023}, and test with hardware in the loop at the Politecnico di Milano MG2Lab microgrid.

\section{Bibliography}\label{bibliography}
